---
title: Why Yield Matters
output: html_document
---



```{r setup, include=FALSE}

########################################
######## INITIALIZING WORKSPACE ########
########################################

library(dplyr)  
library(tidyr)  
library(ggplot2)
library(knitr)
library(xtable)
library(kableExtra)

temp_increase <- 0.01


##############################
######## LOADING DATA ########
##############################

master_data <- read.csv("master_data.csv", stringsAsFactors = FALSE)



###############################
######## MODELING DATA ########
###############################

## Fitting final model to data  

master_data %>%
  glm(formula = ENROLLED ~ poly(APSS_SCORE, 3) + as.factor(ENTRY_YEAR),
      family  = "binomial") -> temp_glm_enroll

master_data %>%
  filter(ENROLLED == 1) %>%
  glm(formula = RETAINED ~ poly(APSS_SCORE, 1) + as.factor(ENTRY_YEAR),
      family  = "binomial") -> temp_glm_retain
  

## Using models to get predictions for data

master_data %>%
  mutate(PRED_ENROLL = predict(object  = temp_glm_enroll, 
                               newdata = ., 
                               type    = "response"),
         PRED_RETAIN = predict(object  = temp_glm_retain, 
                               newdata = ., 
                               type    = "response")) -> master_data
  
  
rm(temp_glm_enroll, temp_glm_retain)




#################################
######## SIMULATING DATA ########
#################################

## Creating the new admit cutoff

master_data %>%
  arrange(ENTRY_YEAR,
          desc(APSS_SCORE)) %>%
  group_by(ENTRY_YEAR) %>%
  mutate(NEW_ADMIT = ifelse(cumsum(PRED_ENROLL + temp_increase) <= sum(ENROLLED), 1, 0)) %>%
  ungroup()  -> master_data



## Running 5,000 simulations of enrollment based on new admit criteria

new_class_stats <- data.frame()


for (i in 1:50) {

  ## Simiulating the new class
  
  master_data %>%
    mutate(temp_enroll = runif(n = nrow(.))) %>%
    filter(NEW_ADMIT    == 1,
          PRED_ENROLL + temp_increase > temp_enroll) -> temp_data
  
  
  ## Summarizing stats for the new class
  
  temp_data %>%
    group_by(ENTRY_YEAR) %>%
    summarise(Interation = i,
              COUNT      = n(),
              APSS_SCORE = mean(APSS_SCORE),
              HS_GPA     = mean(HS_GPA, na.rm = TRUE),
              RETAIN     = mean(PRED_RETAIN)
              ) -> temp_new_class_stats
  
  
  ## Writing data 
  
  new_class_stats <- rbind(new_class_stats, temp_new_class_stats)
  
  rm(temp_data, temp_new_class_stats)
  if(i %% 500 == 0) {gc()}

}
rm(i)


#############################
######## DATA TABLES ########
#############################

## Yield over time

master_data %>%
  group_by(ENTRY_YEAR) %>%
  summarise(YIELD = mean(ENROLLED)) -> table_1


## Class size with increased yield

master_data %>%
  group_by(ENTRY_YEAR) %>%
  summarise(OLD_ENROLL = sum(ENROLLED),
            NEW_ENROLL = round(sum(PRED_ENROLL+temp_increase), 0),
            DIFF       = NEW_ENROLL - OLD_ENROLL) %>%
  ungroup() -> table_2



## Old Class Stats vs New Class Stats

master_data %>%
  group_by(ENTRY_YEAR) %>%
  summarise(APSS_SCORE = mean(APSS_SCORE),
            HS_GPA     = mean(HS_GPA)) %>%
  ungroup() -> temp_old_stats

new_class_stats %>%
  group_by(ENTRY_YEAR) %>%
  summarise(APSS_SCORE = mean(APSS_SCORE),
            HS_GPA     = mean(HS_GPA)) %>%
  ungroup() -> temp_new_stats



####################################
######## DATA VISUALIZATION ########
####################################

## APSS vs Enrolling

master_data %>%
  ggplot() +
    stat_smooth(aes(x = APSS_SCORE,
                    y = ENROLLED),
                method      = "glm",
                formula = y ~ poly(x, 3)) +
    geom_histogram(aes(x = APSS_SCORE,
                       y = ..count../sum(..count..)/.3),
                   binwidth = 1,
                   fill     = "grey30") +
    scale_x_continuous(name = "Admissions Score") +
    scale_y_continuous(name   = "Likelihood of Enrolling",
                       labels = scales::percent) +
    labs(caption = "Figure 1: ") +
    theme(axis.ticks = element_blank(),
          axis.title = element_text(size = 18),
          axis.text  = element_text(size = 14),
          plot.caption    = element_text(size = 15,
                                         hjust = 0)) -> plot_1



## Normal Yield: APSS vs Retain

master_data %>%
  filter(ENROLLED == 1) %>%
  ggplot() +
    stat_smooth(aes(x = APSS_SCORE,
                    y = RETAINED),
                method      = "glm",
                formula = y ~ poly(x, 1)) +
    geom_histogram(aes(x = APSS_SCORE,
                       y = ..count../sum(..count..)/.1),
                   binwidth = 1,
                   fill     = "grey30") +
    scale_x_continuous(name = "Admissions Score") +
    scale_y_continuous(name   = "Likelihood of Retaining",
                       labels = scales::percent) +
    labs(caption = "Figure 2: ") +
    theme(axis.ticks = element_blank(),
          axis.title = element_text(size = 18),
          axis.text  = element_text(size = 14),
          plot.caption    = element_text(size = 15,
                                         hjust = 0)) -> plot_2



## Hypothetical Increased Yield: APSS vs YIELD

master_data %>%
  ggplot() +
    stat_smooth(aes(x = APSS_SCORE,
                    y = ENROLLED),
                method  = "glm",
                formula = y ~ poly(x, 3),
                se      = FALSE,
                linetype = 3) +
    stat_smooth(aes(x = APSS_SCORE,
                    y = ENROLLED),
                method  = "glm",
                formula = y + 0.01 ~ poly(x, 3)) +
    geom_histogram(aes(x    = APSS_SCORE,
                       y    = ..count../sum(..count..)/.3,
                       fill = as.factor(-NEW_ADMIT)),
                   binwidth = 1) +
    scale_x_continuous(name = "Admissions Score") +
    scale_y_continuous(name   = "Likelihood of Enrolling",
                       labels = scales::percent) +
    scale_fill_manual(values = c("grey30", "firebrick2")) +
    labs(caption = strwrap("Figure 3: Effect of increased yield. The red shaded area represents students who could be denied admission without affect overall enrollment if yield increased")) +
    theme(axis.ticks      = element_blank(),
          legend.position = "none",
          axis.title      = element_text(size = 18),
          axis.text       = element_text(size = 14),
          plot.caption    = element_text(size = 15,
                                         hjust = 0)) -> plot_3


```

<div class = "container">

<div class = "row">

### Introduction


At many colleges and universities, faculty are required to assist with undergraduate recruitment. And, this is particulary true at small, tuition-dependant institutions like Eckerd College where faculty often meet individually with prospective students/parents and participate in events organized by the college's admissions office - such as open houses and discussion panels. These obligations can be time consuming and substantially add to the workload of the faculty who are already busy teaching and mentoring students. So most faculty understand the importance of recruitment as it relates to the financial health of the institution, it's not surprising that faculty are sometimes frustrated by these added responsibilities.

Here, we demonstrate the importance of recruitment as it relates to admissions and enrollment management by simulating a `r scales::percent(temp_increase)` increase in yield. 


</div>


<div class = "row">
### Applicant Quality, Yield, and Retention

```{r echo = FALSE, out.width = "50%", out.extra = 'style="float:right; padding:25px"'}

plot_1


```


Each applicant is scored by the admissions office on a 0 to 100 scale. This score is based on the student's high school GPA and standardized test scores and meant to reflect the overall quality of the applicant. Whlie not the only factor considered, decisions regarding admission and financial aid are heavily based on these scores.

Yield varies as a function of applicant quality (Figure 1). In order to encourage high achieving students to enroll, Eckerd College awards more merit-based aid to studunts with high admissions scores. Still, competition for these students is stiff, since all colleges and universities compete against each other to attract the best students. As a result, students with strong GPAs and test scores have many options. 


</div>


<div class = "row">


```{r echo = FALSE, out.width = "50%", out.extra = 'style="float:right; padding:25px"'}

## APSS Points vs Retention

plot_2

```

Simialry, the likelihood that a student is retained through the start of their second year also varies as a function of their admission scores, such that students with higher admissions scores more likely to be retained. 

</div>



<div class = "row">

### Effect of Increased Yield

```{r echo = FALSE, out.width = "50%", out.extra = 'style="float:right; padding:25px"'}


plot_3



```

Natually, if yield increased and the number of admidted students remained the same, the number of student who would enroll would increase - an increase of roughly `r round((sum(master_data$PRED_ENROLL + temp_increase) - sum(master_data$ENROLLED))/length(unique(master_data$ENTRY_YEAR)))` students per year.

However, if the college sought to maintain enrollment levels, a `r scales::percent(temp_increase)` increase in yield would mean that school could deny admission to lowest scoring `r scales::percent(mean(master_data$NEW_ADMIT == 0), 0.1)` of admitted students. 

</div>


### Conclusion



</div>
